---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  cache = FALSE
)
library(rpyANTs)
```

# rpyANTs

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/rpyANTs)](https://CRAN.R-project.org/package=rpyANTs)
[![R-check](https://github.com/dipterix/rpyANTs/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/dipterix/rpyANTs/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

`rpyANTs` is a package that ports `ANTsPy`, a `Python` implementation of `ANTs` into R using R-Python interpreter package `reticulate`. 

> Disclaimer: This is NOT the official [`ANTsR` package](https://github.com/ANTsX/ANTsR).  

## Installation

The installation requires one-line extra setup

```r
# Install from CRAN
install.packages("rpyANTs")

# Install from nightly dev builder
# install.packages("rpyANTs", repos = "https://dipterix.r-universe.dev")


# set up ANTs
rpyANTs::install_ants()
```

`install_ants` creates an isolated `Python` environment managed by [`RAVE`](https://openwetware.org/wiki/RAVE). This environment does not conflict nor affect your existing Python installations.

## How to use

To load `ANTs`

```{r example1, eval = FALSE}
library(rpyANTs)

# Whether ANTs is available
ants_available()

# Load ANTs into R
ants
```

In R, we use `$` to get module functions or class members. For example:

```{r example2}
ants$add_noise_to_image
```

The following R code translates Python code into R:

```{r example3}
# >>> img = ants.image_read(ants.get_ants_data('r16'))
img <- ants$image_read(ants$get_ants_data('r16'))

# >>> noise_image1 = ants.add_noise_to_image(img, 'additivegaussian', (0.0, 1.0))
noise_image1 <- ants$add_noise_to_image(
  img, 'additivegaussian', 
  noise_parameters = tuple(0.0, 1.0)
)

# >>> noise_image2 = ants.add_noise_to_image(img, 'saltandpepper', (0.1, 0.0, 100.0))
noise_image2 <- ants$add_noise_to_image(
  img, 'saltandpepper', 
  noise_parameters = tuple(0.1, 0.0, 100.0)
)

# >>> noise_image3 = ants.add_noise_to_image(img, 'shot', 1.0)
noise_image3 <- ants$add_noise_to_image(
  img, 'shot', 
  noise_parameters = 1.0
)

# >>> noise_image4 = ants.add_noise_to_image(img, 'speckle', 1.0)
noise_image4 <- ants$add_noise_to_image(
  img, 'speckle', 
  noise_parameters = 1.0
)

```

To load imaging data into R

```{r}
orig_array <- py_to_r(img$numpy())
noise_array1 <- py_to_r(noise_image1$numpy())
noise_array2 <- py_to_r(noise_image2$numpy())
noise_array3 <- py_to_r(noise_image3$numpy())
noise_array4 <- py_to_r(noise_image4$numpy())

# plot via R
layout(matrix(c(1,1,2,3,1,1,4,5), nrow = 2, byrow = TRUE))
par(mar = c(0.1, 0.1, 0.1, 0.1), bg = "black", fg = "white")
pal <- grDevices::gray.colors(256, start = 0, end = 1)

image(orig_array, asp = 1, axes = FALSE, 
      col = pal, zlim = c(0, 255))
image(noise_array1, asp = 1, axes = FALSE, 
      col = pal, zlim = c(0, 255))
image(noise_array2, asp = 1, axes = FALSE, 
      col = pal, zlim = c(0, 255))
image(noise_array3, asp = 1, axes = FALSE, 
      col = pal, zlim = c(0, 255))
image(noise_array4, asp = 1, axes = FALSE, 
      col = pal, zlim = c(0, 255))
```

## Upgrade `ANTs`

To upgrade `ANTs`, first update `rpyANTs`, then upgrade `ANTsPyx`

```r
install.packages("rpyANTs")
rpymat::add_packages(packages = "antspyx", pip = TRUE)
```

## Advanced use case

#### Run/Debug `Python` scripts

`rpyANTs` ports functions that allows to run `Python` scripts. For example:


```{r}
library(rpyANTs)

script_path <- tempfile(fileext = ".py")
writeLines(con = script_path, text = r"(

# This is Python script
import ants
print(ants.__version__)

)")

run_script(script_path)
```

You can also run `Python` interactive in R (yes, you are correct). Simply run

```r
rpyANTs::repl_python()
```

The console prefix will change from `>` to `>>>`, meaning you are in `Python` mode:

```
> rpyANTs::repl_python()
Python 3.8.16 (/Users/dipterix/Library/r-rpymat/miniconda/envs/rpymat-conda-env/bin/python3.8)
Reticulate 1.26 REPL -- A Python interpreter in R.
Enter 'exit' or 'quit' to exit the REPL and return to R.
>>> 
```

Try some Python code!

```
>>> import ants
>>> help(ants.registration)
```

To exit Python mode, type `exit` (no parenthesis) and hit enter key

```
>>> exit
> 
```


#### Data conversions

Native R variables can be easily converted to `Python` and back via `r_to_py` and `py_to_r`.

For example

```{r}
# R to Python
r_to_py(1)
r_to_py(1L)

# Python to R
py_obj <- py_list(1:3)
class(py_obj)  # <- this is a python object

py_to_r(py_obj)
```

You can also use variables created in R from Python or vice versa:

In the following example, an R object `object_r` is created. In Python, it can be accessed (read-only) via `r.object_r`


```
> object_r <- c(1,2,3)
> repl_python()
Python 3.8.16 (/Users/dipterix/Library/r-rpymat/miniconda/envs/rpymat-conda-env/bin/python3.8)
Reticulate 1.26 REPL -- A Python interpreter in R.
Enter 'exit' or 'quit' to exit the REPL and return to R.
>>> r.object_r
[1.0, 2.0, 3.0]
```

Similarly, a Python object `object_py` is created, and it can be read from `py$object_py`:

```
>>> import numpy as np
>>> object_py = np.array([2,3,4])
>>> exit
> py$object_py
[1] 2 3 4
```

## Known issues

#### Variable types

R is not a type-rigid language. Some functions in `ANTsPy` require specific variable types that are often vague in R. For example the `dimension` argument in function `ants$create_ants_transform` needs to be an integer, but R's default numerical values are `double`. In this case, variable formats need to be explicitly given.

Here are several examples

1. Explicit integers

```r
# ants$create_ants_transform(dimension = 3)     # <- error
ants$create_ants_transform(dimension = 3L)      # < XXXL is an explicit integer
```

2. `Tuple`, `list`, and `dictionary`

A Python `tuple` is a vector that cannot alter lengths. 

```r
# Wrong as `aff_iterations` needs to be a tuple
# ants$registration(fixed, moving, ..., aff_iterations = c(6L, 4L, 2L, 1L))

ants$registration(fixed, moving, ..., aff_iterations = tuple(6L, 4L, 2L, 1L))
```

Similar conversions can be done via `py_list`, `py_dict`.

#### Operators

Currently there is no operator support in `rpyANTs`. For example, `ANTsImage` has Python operators:

```python
# Python code
import ants
import numpy as np
image = ants.image_read(ants.get_ants_data('r16'))

ants.plot(image > 10)
```


The operator is not available in R (don't worry, I will add it in the future, it takes time, OK?), hence it takes a little bit work-around.

Work-around version 1: call operators directly

```r
library(rpyANTs)
image <- ants$image_read(ants$get_ants_data('r16'))

# cannot use image > 10 ***for now***
threshold <- image$`__gt__`(10)
ants$plot(threshold)
```

Work-around version 2: If you don't know how Python operators work, use Python directly

```r
library(rpyANTs)
image <- ants$image_read(ants$get_ants_data('r16'))

# Create an R variable in Python!
py_run_string("r.threshold = r.image > 10", local = TRUE, convert = FALSE)
ants$plot(threshold)
```

## Citation

This is a general citation:

> Avants, B.B., Tustison, N. and Song, G., 2009. Advanced normalization tools (ANTS). The Insight Journal, 2(365), pp.1-35.

Please check their official website to cite specific methods.

## License

This package `rpyANTs` is released under Apache-2.0 license (Copyright: Zhengjia Wang). 
The underlying `ANTsPy` is released under Apache-2.0 license (Copyright: ANTs contributors). 

